---
name: Bryan M. Maitland
position: "Research Fish Biologist"
address: "USDA Forest Service, Rocky Mountain\\tabularnewline Research Station, Boise, ID, USA"
email: "bmaitland101@gmail.com"
orcid: 0000-0002-4491-5064
linkedin: bryanmaitland
github: bmait101
date: "`r format(Sys.time(), '%B %Y')`"
output: 
  vitae::hyndman:
    pandoc_args: ["--lua-filter=bold-surname.lua"]
header-includes:
  \definecolor{headcolor}{HTML}{000088}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  warning = FALSE, 
  message = FALSE
  )

library(vitae)
library(dplyr)
library(RefManageR)
source(here::here("R/pretty.R"))
source(here::here("R/rpackages.R"))
source(here::here("R/gcite.R"))
# source(here::here("R/add_bib_section.R"))
```

# Employment

```{r employment, results='asis'}
tribble(
  ~Start, ~End, ~Position,
  2023, 2024, "\\textbf{Research Associate}, Department of Civil and Environmental Engineering, University of Wisconsin-Madison",
  2024, NA, "\\textbf{Research Fish Biologist}, Rocky Mountain Research Station, USDA Forest Service",
  # 2015, 2020, "\\textbf{Graduate Research Assistant}, Department of Zoology and Physiology, University of Wyoming",
  # 2013, 2015, "\\textbf{Graduate Research Assistant}, Department of Renewable Resources, University of Alberta",
  # 2014, 2014, "\\textbf{Research Technician}, Athabasca River Oil Sands Project, University of Alberta",
  # 2013, 2014, "\\textbf{Research Technician}, Aquatics Division, Parks Canada",
  # 2011, 2013, "\\textbf{Fisheries Research Technician}, USGS Great Lakes Science Center",
  # 2010, 2011, "\\textbf{Laboratory Assistant}, Department of Environmental and Forest Biology, SUNY: ESF",
  # 2010, 2011, "\\textbf{Research Technician}, Thousand Islands Biological Station, SUNY: ESF",
  ) |>
  arrange(-Start) |>
  mutate(
    Start = as.character(Start),
    End = as.character(End),
    End = tidyr::replace_na(End, ""),
    Years = if_else(End == Start,
      as.character(Start),
      paste(Start, "--", End, sep = ""),
      paste(Start, "--", End, sep = "")
    )
  ) |>
  select(Years, Position) |>
  baretable()
```

# Education

```{r degrees, results='asis'}
tribble(
  ~ Degree, ~ Year, ~ Institution,
  "EPA Postdoctoral Fellow", "2022-2023", "University of Wisconsin-Madison",
  "Water Resources Fellow", "2020-2022", "University of Wisconsin-Madison",
  "Ph.D. Ecology", "2020", "University of Wyoming",
  "M.Sc. Conservation Biology", "2015", "University of Alberta",
  "B.S. Conservation Biology", "2011", "SUNY: College of Environmental Science and Forestry"
  ) |> 
  baretable()

# tribble(
  # ~ Degree, ~ Year, ~ Institution, ~ Where,
  # "EPA Postdoctoral Fellow", "2022-2023", "University of Wisconsin-Madison", "Madison, WI, USA",
  # "Water Resources Postdoctoral Fellow", "2020-2022", "University of Wisconsin-Madison", "Madison, WI, USA",
  # "Ph.D. Ecology", "2020", "University of Wyoming", "Laramie, WY, USA",
  # "M.Sc. Conservation Biology", "2015", "University of Alberta", "Edmonton, AB, CAN",
  # "B.S. Conservation Biology", "2011", "SUNY: ESF", "Syracuse, NY, USA"
# ) %>% 
#   detailed_entries(Degree, Year, Institution, Where)
```

# Awards and honours

```{r awards, results='asis'}
tribble(
  ~Year, ~Award,
  2024, "Outstanding Service Award, USDA FS Rocky Mountain Research Station",
  2021, "Best Professional Presentation Award, Wisconsin Chapter American Fisheries Society",
  2019, "Dean’s Graduate Scholar Award, College of Arts and Sciences University of Wyoming",
  2018, "Eugene Maughan Graduate Award, Western Division American Fisheries Society",
  2018, "Hank Gardner Graduate Award, Department of Zoology and Physiology University of Wyoming",
  2017, "Clemens-Rigler Award, Canadian Aquatic Resources Section American Fisheries Society",
  2017, "Board of Visitors Award, College of Arts and Sciences University of Wyoming",
  2017, "Dennis Anderson Fisheries Award, Department Zoology and Physiology University of Wyoming ",
  2017, "William Trachtenberg Scholarship, Sustainable Fisheries Foundation",
  2017, "Vern Bressler Fisheries Award, Department of Zoology and Physiology University of Wyoming",
  2015, "Excellence in Fisheries Award, Canadian Aquatic Resources Section American Fisheries Society",
  2015, "Clemens-Rigler Award, Canadian Aquatic Resources Section American Fisheries Society"
  ) |>
  baretable()
```

# Research

```{r article-totals}
pubs <- RefManageR::ReadBib("refs_articles.bib", check = FALSE) |> 
  as_tibble() |> 
  filter(bibtype == "Article")
```

*   Since 2015 I have authored `r NROW(pubs)` papers in peer-reviewed journals on fisheries, conservation, and food web topics.
*   On Google Scholar my h-index is `r bmm_cite$All[2]` with total citations of `r prettyNum(bmm_cite$All[1], big.mark=",")` (as at `r format(Sys.time(), "%e %B %Y")`).
*   I have coauthored `r NROW(bmm_pkgs)` R package as a result of my research. There have been `r round(sum(bmm_pkgs$downloads, na.rm=TRUE))` downloads of my packages.

# Grants

```{r grants}
grants <- readr::read_csv("grant_income.csv")

sumgrants <- grants |> 
  filter(Status %in% c("Complete", "In progress")) |> 
  pull(Amount) |> 
  sum(na.rm = TRUE)
```

*   I have acquired `r scales::dollar_format()(sumgrants)` in research grants since 2015.\vspace*{0.2cm}\enlargethispage*{0.5cm}

```{r grants-table, results='asis'}
grants |>
  # filter(Amount > 100000, Type == "Research") |>
  filter(Status %in% c("Complete", "In progress", "Pending")) |> 
  mutate(
    Time = paste0(Start, ifelse(is.na(End), "", paste0("--", End))),
    Entry = paste(People, ". ``", Title, "''. ",
      "\\emph{Funding from ", Source, "}",
      sep = ""
    ),
    Amount = dollars(Amount)
  ) |>
  arrange(desc(Time), desc(Amount)) |>
  select(Time, Entry, Amount) %>%
  baretable(
    align = "llp{13.3cm}r",
    hline.after = c(0, NROW(.)),
    size = "\\fontsize{10}{12}",
    add.to.row = list(
      pos = as.list(seq(NROW(.))),
      command = rep("\\\\[-0.32cm]", NROW(.))
    )
  )
```

<!-- \vspace*{0.2cm}\enlargethispage*{0.5cm} -->

<!-- ## Invited, not awarded -->

<!-- Below are invited full proposals that were not awarded: \vspace*{0.2cm}\enlargethispage*{0.5cm} -->

```{r grantstable-invite, results='asis', eval=FALSE}
grants |>
  filter(!Status %in% c("Complete", "In progress", "Pending")) |> 
  mutate(
    Time = paste0(Start, ifelse(is.na(End), "", paste0("--", End))),
    Entry = paste(People, ". ``", Title, "''. ",
      "\\emph{Funding from ", Source, "}",
      sep = ""
    ),
    Amount = dollars(Amount)
  ) |>
  arrange(desc(Time), desc(Amount)) |>
  select(Time, Entry, Amount) %>%
  baretable(
    align = "llp{13.3cm}r",
    hline.after = c(0, NROW(.)),
    size = "\\fontsize{10}{12}",
    add.to.row = list(
      pos = as.list(seq(NROW(.))),
      command = rep("\\\\[-0.32cm]", NROW(.))
    )
  )
```

\newpage

# Publications (chronological)

## Theses

```{r}
bibliography_entries("refs_theses.bib") |> 
  arrange(issued)
```

## Pre-prints and in-review

```{r}
bibliography_entries("refs_articles.bib") |> 
  filter(type == "manuscript") |> 
  arrange(issued, desc(author$family))
```

## Refereed research papers

```{r}
bibliography_entries("refs_articles.bib") |> 
  filter(type == "article-journal") |> 
  arrange(issued, desc(author$family))
```

## Technical Reports

```{r}
bibliography_entries("refs_reports.bib") %>%
  arrange(issued, desc(author$family))
```

## Unpublished working papers

```{r}
bibliography_entries("refs_working.bib") %>%
  arrange(issued, desc(author$family))
```

## Software

```{r}
bibliography_entries("refs_pkgs.bib")
```

\newpage

# Specialized Skills and Certifications

## Field Skills

```{r, results='asis'}
tribble(
  ~area, ~skills,
  "Fish Sampling", "Electrofishing, netting, seining, trawling, angling",
  "Aquatic Ecology", "Water quality sampling, invertebrate sampling, aquatic vegetation sampling, habitat assessment",
  "Identification", "Fish (freshwater and marine), amphibians, reptiles, aquatic invertebrates, aquatic plants",
  "Operations", "Piloting and trailering small vessels, rafting, helicopters experience, orienteering",
  "Misc", "Light machinery operation, small power equipment repair, minor electrical repair, handiwork",
  ) |> 
  baretable()
```

## Lab Skills

```{r, results='asis'}
tribble(
  ~area, ~skills,
  "Fisheries", "Fish aging (scales, otoliths), fish dissection, fish tissue processing",
  "Aquatic Ecology", "Water quality analysis (nutrients, chlorophyll a, turbidity), invertebrate processing, aquatic vegetation",
  "Stable Isotopes", "Sample preparation, ligh (C, N, O, S) and heavy (Sr) stable isotope analysis",
  "Molecular", "DNA extraction, PCR, qPCR",
  ) |> 
  baretable()
```

## Computer Skills

```{r, results='asis'}
tribble(
  ~area, ~skills,
  "Operating Systems", "Linux, MacOS, Windows",
  "Programming Languages", "R, Python, BASH, JavaScript",
  "Markup Languages", "Markdown, RMarkdown, CSS, HTML, LaTeX",
  "Data Interchange Formats", "CSV, JSON, XML",
  "Other Languages", "YAML",
  "Version Control", "Git",
  "Text Editors", "RStudio, SQL Server Management Studio, Visual Studio, Visual Studio Code",
  "Microsoft Office", "Excel, Outlook, OneNote, PowerPoint, Word"
  ) |> 
  baretable()
```

## Certifications

```{r, results='asis'}
tribble(
  ~Year, ~Certification, ~Institution,
  2025, "Wilderness First Aid", "Wilderness Medicine Institude",
  2025, "Advanced CPR and First Aid", "American Red Cross",
  2014, "Defensive Driving", "USDA Forest Service",
  2018, "Avalanche Safety Level 1", "National Ski Patrol",
  2013, "Electrofishing Safety", "University of Alberta",
  2013, "MOCC Federal Motor Boat Operator", "US Department of the Interior",
  ) |> 
  baretable()
```

# Professional Service

## Ad-hoc reviewer

```{r reviews1}
n_reviewed <- 49
n_journals <- 15
n_per_yr <- round(n_reviewed / (2025 - 2016), 1)
```

*   I have served as an ad-hoc reviewer for `r n_reviewed` manuscripts from `r n_journals` journals since 2016 ($\sim$ `r n_per_yr` per year). 

## Other peer review

```{r reviews2, results='asis'}
tribble(
  ~Year, ~Details,
  2025, "Lake Champlain Fisheries Research Program, full proposal review",
  2023, "NSF, Division of Environmental Biology, CAREER, full proposal review",
  ) |>
  baretable()
```

## Society membership

```{r societies, results='asis'}
tribble(
  ~Start, ~End, ~Position,
  2010, NA, "American Fisheries Society",
  2010, NA, "Society for Conservation Biology",
  2015, NA, "Ecological Society of America",
  2016, NA, "Society for Freshwater Science",
  2019, NA, "Galápagos Conservancy",
  2016, 2019, "Association for the Science of Limnology and Oceanography",
  2010, 2016, "International Association of Great Lakes Research",
  ) |>
  # arrange(-Start) |>
  mutate(
    Start = as.character(Start),
    End = as.character(End),
    End = tidyr::replace_na(End, ""),
    Years = if_else(End == Start,
      as.character(Start),
      paste(Start, "--", End, sep = ""),
      paste(Start, "--", End, sep = "")
    )
  ) |>
  select(Years, Position) |>
  baretable()
```

## Society service

```{r society-service, results='asis'}
tribble(
  ~Start, ~End, ~Position,
  2015, NA, "\\textbf{Member}, Freshwater Working Group, Society for Conservation Biology",
  2021, 2022, "\\textbf{Member}, Fisheries Working Group, Wisconsin Initiative on Climate Change Impacts",
  2020, 2022, "\\textbf{Member}, Trout Species Team, Wisconsin Department of Natural Resources",
  2016, 2017, "\\textbf{President}, UWyo Student Chapter, American Fisheries Society",
  2016, 2018, "\\textbf{Excom member}, CO-WY State Chapter, American Fisheries Society",
  ) |>
  # arrange(-Start) |>
  mutate(
    Start = as.character(Start),
    End = as.character(End),
    End = tidyr::replace_na(End, ""),
    Years = if_else(End == Start,
      as.character(Start),
      paste(Start, "--", End, sep = ""),
      paste(Start, "--", End, sep = "")
    )
  ) |>
  select(Years, Position) |>
  baretable()
```

## Conference organization and service

```{r conf-service, results='asis'}
tribble(
  ~Year, ~Details,
  2024, "Abstract reviewer, American Geophysical Union, Fall Meeting",
  ) |>
  baretable()
```

# University service

## Committees, Groups, and Councils

```{r uni-serivce, results='asis'}
tribble(
  ~Start, ~End, ~Position,
  2022, 2024, "Field Operation Facilitation and Boat Committee, Center for Limnology, UW-Madison",
  2019, 2020, "Faculty Search Committee, Department of Zoology and Physiology, University of Wyoming",
  2016, 2020, "Social Media Coordinator, Program in Ecology, University of Wyoming",
  2016, 2017, "Graduate Council, student representative, University of Wyoming",
  2016, 2017, "President, Program in Ecology, University of Wyoming",
  ) |>
  # arrange(-Start) |>
  mutate(
    Start = as.character(Start),
    End = as.character(End),
    End = tidyr::replace_na(End, ""),
    Years = if_else(End == Start,
      as.character(Start),
      paste(Start, "--", End, sep = ""),
      paste(Start, "--", End, sep = "")
    )
  ) |>
  select(Years, Position) |>
  baretable()
```

## Teching

```{r teaching, results='asis'}
tribble(
  ~ Year, ~ Course, ~ Role, ~ Institution,
  "2020", "ZOO 4330 - Ichthyology", "Instructor", "University of Wyoming",
  "2019", "ZOO 4330 - Ichthyology", "Lab Instructor", "University of Wyoming",
  "2018", "ZOO 4330 - Ichthyology", "Lab Instructor", "University of Wyoming",
  "2017", "ZOO 4330 - Ichthyology", "Lab Instructor", "University of Wyoming",
  "2016", "ZOO 4330 - Ichthyology", "Lab Instructor", "University of Wyoming",
  "2015", "ZOO 4310 - Fisheries Management", "Lab Instructor", "University of Wyoming",
  "2014", "RENR 460 - Fisheries Management", "Lab Instructor", "University of Alberta",
  "2013", "RENR 460 - Fisheries Management", "Lab Instructor", "University of Alberta",
  ) |> 
  baretable()
```

## Mentoring

### Undergraduate Students (Research Assistants, lab and field)

```{r mentoring, results='asis'}
tribble(
  ~ Start, ~ End, ~ Name, ~ Institution,
  2023, 2023, "Jonathan Martin", "University of Wisconsin-Madison",
  2018, 2020, "Nathan Barrus", "University of Wyoming",
  2019, 2020, "Russell Brubaker", "University of Wyoming",
  2018, 2019, "Reilly Davis", "University of Wyoming",
  2017, 2017, "Brynn Hirschman", "University of Wyoming",
  2017, 2017, "Austin Jacobson", "University of Wyoming",
  2016, 2016, "Shaun Helsper", "University of Wyoming",
  2016, 2017, "Travis Allison", "University of Wyoming",
  2016, 2016, "Andrew Annear", "University of Wyoming",
  2015, 2015, "Jake Werner", "University of Wyoming",
  ) |>
  baretable()
```

# Selected Media

* Milwaukee Journal Sentinel, Jan 2023. ["Climate change is making Wisconsin trout more vulnerable to floods and heat"](https://www.jsonline.com/story/news/2023/01/25/climate-change-is-making-wisconsin-trout-more-vulnerable-study-shows/69800081007/). (Also covered in separate stories on [WPR Public Radio](https://www.wpr.org/shows/morning-show/miss-wisconsin-crowned-miss-america-trout-feeling-affects-climate-change))
* Water Blogged, Dec 2022. ["A Tale of Two Fishes: Study Maps the Future of Trout Populations in Wisconsin"](https://blog.limnology.wisc.edu/2022/12/19/a-tale-of-two-fishes-study-maps-the-future-of-trout-populations-in-wisconsin/)
* UW-Madison News, Dec 2022. ["Study maps uneasy future of Wisconsin trout populations"](https://news.wisc.edu/study-maps-uneasy-future-of-wisconsin-trout-populations/)
* The Tributary, Nov 2017. [“Stuck in a rut in Anaconda? Crash a cutthroat spawning party!”](https://wdafs.org/download/Vol-41-No-02.pdf/)

# Community Service and Outreach

```{r outreach, results='asis'}
tribble(
  ~ Start, ~ End, ~ Role, ~ Institution,
  2022, 2023, "Food service volunteer", "Ronald McDonald House, Madison, WI",
  2016, 2020, "Science outreach and stream ecology field trips", "Laramie High School, Laramie, WY",
  2016, 2019, "Group leader, Wyoming Bio-Blitz", "Biodiversity Institute, University of Wyoming",
  2016, 2018, "Big Brother", "Big Brothers Big Sisters of Wyoming",
  2017, 2018, "Habitat restoration projects", "Girl Scouts of America",
  2016, 2017, "Science judge", "Wyoming State Science Fair"
  ) |>
  # arrange(-Start) |>
  mutate(
    Start = as.character(Start),
    End = as.character(End),
    End = tidyr::replace_na(End, ""),
    Years = if_else(End == Start,
      as.character(Start),
      paste(Start, "--", End, sep = ""),
      paste(Start, "--", End, sep = "")
    )
  ) |>
  select(Years, Role, Institution) |>
  baretable()
```

