---
name: Bryan M. Maitland, Ph.D.
position: "Research Fish Biologist"
address: "Rocky Mountain Research Station,\\tabularnewline USDA Forest Service, Boise, ID, USA"
email: "bmaitland101@gmail.com"
orcid: 0000-0002-4491-5064
linkedin: bryanmaitland
github: bmait101
date: "`r format(Sys.time(), '%B %Y')`"
output: 
  vitae::hyndman:
    pandoc_args: ["--lua-filter=bold-surname.lua"]
header-includes:
  \definecolor{headcolor}{HTML}{000088}
---

\newcounter{bib}

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  warning = FALSE, 
  message = FALSE
  )

library(vitae)
library(dplyr)
library(RefManageR)
source(here::here("R/pretty.R"))
source(here::here("R/rpackages.R"))
source(here::here("R/gcite.R"))
# source(here::here("R/add_bib_section.R"))
```

# Employment

```{r employment, results='asis'}
tribble(
  ~Start, ~End, ~Position,
  2024, NA, "\\textbf{Research Fish Biologist}, Rocky Mountain Research Station, USDA Forest Service",
  2023, 2024, "\\textbf{Research Associate}, Dept. of Civil and Environmental Engineering, University of Wisconsin-Madison",
  2022, 2023, "\\textbf{EPA Postdoctoral Fellow}, Center for Limnology, University of Wisconsin-Madison",
  2020, 2022, "\\textbf{Water Resources Postdoctoral Fellow}, WI Water Resrouces Institute, University of Wisconsin-Madison",
  # 2015, 2020, "\\textbf{Graduate Research Assistant}, Department of Zoology and Physiology, University of Wyoming",
  # 2013, 2015, "\\textbf{Graduate Research Assistant}, Department of Renewable Resources, University of Alberta",
  # 2014, 2014, "\\textbf{Research Technician}, Athabasca River Oil Sands Project, University of Alberta",
  # 2013, 2014, "\\textbf{Research Technician}, Aquatics Division, Parks Canada",
  # 2011, 2013, "\\textbf{Fisheries Research Technician}, USGS Great Lakes Science Center",
  # 2010, 2011, "\\textbf{Laboratory Assistant}, Department of Environmental and Forest Biology, SUNY: ESF",
  # 2010, 2011, "\\textbf{Research Technician}, Thousand Islands Biological Station, SUNY: ESF",
  ) |>
  arrange(-Start) |>
  mutate(
    Start = as.character(Start),
    End = as.character(End),
    End = tidyr::replace_na(End, ""),
    Years = if_else(End == Start,
      as.character(Start),
      paste(Start, "--", End, sep = ""),
      paste(Start, "--", End, sep = "")
    )
  ) |>
  select(Years, Position) |>
  baretable()
```

# Education

```{r degrees, results='asis'}
tribble(
  ~ Degree, ~ Year, ~ Feild, ~ Institution,
  "Ph.D.", 2020, "Ecology", "University of Wyoming",  
  "M.Sc." ,2015, "Conservation Biology", "University of Alberta", 
  "B.S.", 2011, "Conservation Biology \\textit{cum luade}", "SUNY: College of Environmental Science and Forestry",
  ) |> 
  baretable()
```

# Awards and honours

```{r awards, results='asis'}
tribble(
  ~Year, ~Award,
  2024, "Outstanding Service Award, USDA FS Rocky Mountain Research Station",
  2021, "Best Professional Presentation Award, Wisconsin Chapter American Fisheries Society",
  2019, "Dean’s Graduate Scholar Award, College of Arts and Sciences University of Wyoming",
  2018, "Eugene Maughan Graduate Award, Western Division American Fisheries Society",
  2018, "Hank Gardner Graduate Award, Department of Zoology and Physiology University of Wyoming",
  2017, "Clemens-Rigler Award, Canadian Aquatic Resources Section American Fisheries Society",
  2017, "Board of Visitors Award, College of Arts and Sciences University of Wyoming",
  # 2017, "Dennis Anderson Fisheries Award, Department Zoology and Physiology University of Wyoming ",
  2017, "William Trachtenberg Scholarship, Sustainable Fisheries Foundation",
  2017, "Vern Bressler Fisheries Award, Department of Zoology and Physiology University of Wyoming",
  2015, "Excellence in Fisheries Award, Canadian Aquatic Resources Section American Fisheries Society",
  2015, "Clemens-Rigler Award, Canadian Aquatic Resources Section American Fisheries Society"
  ) |>
  baretable()
```

# Research

```{r article-totals}
n_pubs <- RefManageR::ReadBib("refs_articles.bib", check = FALSE) |> 
  as_tibble() |> 
  filter(bibtype == "Article") |> 
  nrow()

n_talks <- RefManageR::ReadBib("refs_presentations.bib", check = FALSE) |> 
  as_tibble() |> 
  nrow()
```

* My current research focuses on the ecology, management, and conservation of freshwater fishes, with a particular emphasis on coldwater fishes in the Rocky Mountains, Pacific Northwest, and Great Lakes regions.
* Since 2015 I have authored `r n_pubs` papers in peer-reviewed journals on fisheries and conservation (pages \pageref{papersstart}--\pageref{papersend}).
* On Google Scholar my h-index is `r bmm_cite$All[2]` with total citations of `r prettyNum(bmm_cite$All[1], big.mark=",")` (as at `r format(Sys.time(), "%e %B %Y")`).
* I have coauthored `r NROW(bmm_pkgs)` R package as a result of my research with a total of `r round(sum(bmm_pkgs$downloads, na.rm=TRUE))` downloads (page \pageref{packagestart}).
* Since 2011 I have given or co-authored `r n_talks` presentations on my research (pages \pageref{presstart}--\pageref{presend}).

# Grants

```{r grants}
grants <- readr::read_csv("grant_income.csv")

sumgrants <- grants |> 
  filter(Status %in% c("Complete", "In progress")) |> 
  pull(Amount) |> 
  sum(na.rm = TRUE)
```

*   I have acquired `r scales::dollar_format()(sumgrants)` in research grants since 2015.\vspace*{0.2cm}\enlargethispage*{0.5cm}

```{r grants-table, results='asis'}
grants |>
  # filter(Amount > 100000, Type == "Research") |>
  filter(Status %in% c("Complete", "In progress", "Pending")) |> 
  mutate(
    Time = paste0(Start, ifelse(is.na(End), "", paste0("--", End))),
    Entry = paste(People, ". ``", Title, "''. ",
      "\\emph{Funding from ", Source, "}",
      sep = ""
    ),
    Amount = dollars(Amount)
  ) |>
  arrange(desc(Time), desc(Amount)) |>
  select(Time, Entry, Amount) %>%
  baretable(
    align = "llp{13.3cm}r",
    hline.after = c(0, NROW(.)),
    size = "\\fontsize{10}{12}",
    add.to.row = list(
      pos = as.list(seq(NROW(.))),
      command = rep("\\\\[-0.32cm]", NROW(.))
    )
  )
```

<!-- \vspace*{0.2cm}\enlargethispage*{0.5cm} -->

<!-- ## Invited, not awarded -->

<!-- Below are invited full proposals that were not awarded: \vspace*{0.2cm}\enlargethispage*{0.5cm} -->

```{r grantstable-invite, results='asis', eval=FALSE}
grants |>
  filter(!Status %in% c("Complete", "In progress", "Pending")) |> 
  mutate(
    Time = paste0(Start, ifelse(is.na(End), "", paste0("--", End))),
    Entry = paste(People, ". ``", Title, "''. ",
      "\\emph{Funding from ", Source, "}",
      sep = ""
    ),
    Amount = dollars(Amount)
  ) |>
  arrange(desc(Time), desc(Amount)) |>
  select(Time, Entry, Amount) %>%
  baretable(
    align = "llp{13.3cm}r",
    hline.after = c(0, NROW(.)),
    size = "\\fontsize{10}{12}",
    add.to.row = list(
      pos = as.list(seq(NROW(.))),
      command = rep("\\\\[-0.32cm]", NROW(.))
    )
  )
```

\newpage

# Publications (chronological within subgroups)

## Theses

```{r theses}
bibliography_entries("refs_theses.bib") |> 
  arrange(issued)
```

\refstepcounter{bib}\label{papersstart}

## Refereed research papers

```{r articles}
bibliography_entries("refs_articles.bib") |> 
  filter(type == "article-journal") |> 
  arrange(issued, desc(author$family))
```

\refstepcounter{bib}\label{papersend}

## Technical Reports

```{r techreports}
bibliography_entries("refs_reports.bib") %>%
  arrange(issued, desc(author$family))
```

## Pre-prints and in-review

```{r preprints}
bibliography_entries("refs_articles.bib") |> 
  filter(type == "manuscript") |> 
  arrange(issued, desc(author$family))
```

## Unpublished working papers

```{r unpublished}
bibliography_entries("refs_working.bib") %>%
  arrange(issued, desc(author$family))
```

## Software

\refstepcounter{bib}\label{packagestart}

```{r software}
bibliography_entries("refs_pkgs.bib")
```

\refstepcounter{bib}\label{packageend}

\newpage

# Specialized Skills and Certifications

## Field Skills

```{r, results='asis'}
tribble(
  ~area, ~skills,
  "Fish Sampling", "Electrofishing, netting, seining, trawling, angling",
  "Aquatic Ecology", "Water quality sampling, invertebrate sampling, aquatic vegetation sampling, habitat assessment",
  "Identification", "Fish (freshwater and marine), amphibians, reptiles, aquatic invertebrates, aquatic plants",
  "Operations", "Piloting and trailering small vessels, rafting, helicopters experience, orienteering",
  "Other", "Light machinery operation, small power equipment repair, minor electrical repair, handiwork",
  ) |> 
  baretable()
```

## Lab Skills

```{r, results='asis'}
tribble(
  ~area, ~skills,
  "Fisheries", "Fish aging (scales, otoliths), fish dissection, fish tissue processing",
  "Aquatic Ecology", "Water quality analysis (nutrients, chlorophyll a, turbidity), invertebrate processing, aquatic vegetation",
  "Stable Isotopes", "Sample preparation, ligh (C, N, O, S) and heavy (Sr) stable isotope analysis",
  "Molecular", "DNA extraction, PCR, qPCR",
  ) |> 
  baretable()
```

## Computer Skills

```{r, results='asis'}
tribble(
  ~area, ~skills,
  "Operating Systems", "Linux, MacOS, Windows",
  "Programming Languages", "R, Python, BASH, JavaScript",
  "Markup Languages", "Markdown, RMarkdown, CSS, HTML, LaTeX",
  "Data Interchange Formats", "CSV, JSON, XML",
  "Other Languages", "YAML",
  "Version Control", "Git",
  "Text Editors", "RStudio, SQL Server Management Studio, Visual Studio, Visual Studio Code",
  "Microsoft Office", "Excel, Outlook, OneNote, PowerPoint, Word"
  ) |> 
  baretable()
```

## Analytical Skills

```{r, results='asis'}
tribble(
  ~area, ~skills,
  "Statistical Analysis", "Generalized linear models, mixed effects models, Bayesian statistics, multivariate statistics",
  "Data Visualization", "ggplot2, base R graphics, leaflet, plotly",
  "Data Management", "Data wrangling and cleaning, data management with SQL",
  "Spatial Analysis", "R, QGIS, ArcGIS, Google Earth Engine",
  "Package Development", "R package development, documentation, testing, and maintenance",
  ) |> 
  baretable()
```

## Leadership Development

```{r leadership, results='asis'}
tribble(
  ~Year, ~Program, ~Institution,
  2021, "Equity and Inclusion Workshop", "EQT By Design via WI Sea Grant",
  2008, "Leave No Trace Master Educator", "Center for Outdoor Ethics, NOLS",
  ) |> 
  baretable()
```

## Certifications

```{r, results='asis'}
tribble(
  ~Year, ~Certification, ~Institution,
  2025, "Wilderness First Aid", "Wilderness Medicine Institude",
  2025, "Advanced CPR and First Aid", "American Red Cross",
  2014, "Defensive Driving", "USDA Forest Service",
  2018, "Avalanche Safety Level 1", "National Ski Patrol",
  2013, "Electrofishing Safety", "University of Alberta",
  2013, "MOCC Federal Motor Boat Operator", "US Department of the Interior",
  ) |> 
  baretable()
```

## Additional Academic Training

```{r, results='asis'}
tribble(
  ~Year, ~Course, ~Credits, ~Institution,
  2008, "Environmental Ethics", "2 credits", "University of Utah",
  2008, "Group Leadership Techniques", "2 credits", "University of Utah",
  2008, "Wilderness Skills Practicum, Backpacking", "2 credits", "University of Utah",
  2007, "Wilderness Skills Practicum, River Travel", "2 credits", "University of Utah",
  ) |> 
  baretable()
```

# Society and Professional Service

## Ad-hoc reviewer

```{r reviews1}
n_reviewed <- 49
n_journals <- 15
n_per_yr <- round(n_reviewed / (2025 - 2016), 1)
```

*   I have served as an ad-hoc reviewer for `r n_reviewed` manuscripts from `r n_journals` journals since 2016 ($\sim$ `r n_per_yr` per year). 
*   2025, Lake Champlain Fisheries Research Program, full proposal review
*   2023, NSF, Division of Environmental Biology, CAREER, full proposal review

## Society memberships

```{r societies, results='asis'}
tribble(
  ~Start, ~End, ~Position,
  2010, NA, "American Fisheries Society",
  2010, NA, "Society for Conservation Biology",
  2015, NA, "Ecological Society of America",
  2016, NA, "Society for Freshwater Science",
  2019, NA, "Galápagos Conservancy",
  2016, 2019, "Association for the Science of Limnology and Oceanography",
  2010, 2016, "International Association of Great Lakes Research",
  ) |>
  arrange(Start) |>
  mutate(
    Start = as.character(Start),
    End = as.character(End),
    End = tidyr::replace_na(End, ""),
    Years = if_else(End == Start,
      as.character(Start),
      paste(Start, "--", End, sep = ""),
      paste(Start, "--", End, sep = "")
    )
  ) |>
  select(Years, Position) |>
  baretable()
```

## Society leadership

```{r society-service, results='asis'}
tribble(
  ~Start, ~End, ~Position,
  2015, NA, "Member, Freshwater Working Group, Society for Conservation Biology",
  2021, 2022, "Member, Fisheries Working Group, Wisconsin Initiative on Climate Change Impacts",
  2020, 2022, "Member, Trout Species Team, Wisconsin Department of Natural Resources",
  2016, 2017, "President, UWyo Student Chapter, American Fisheries Society",
  2016, 2018, "Excom Member, CO-WY State Chapter, American Fisheries Society",
  ) |>
  arrange(Start) |>
  mutate(
    Start = as.character(Start),
    End = as.character(End),
    End = tidyr::replace_na(End, ""),
    Years = if_else(End == Start,
      as.character(Start),
      paste(Start, "--", End, sep = ""),
      paste(Start, "--", End, sep = "")
    )
  ) |>
  select(Years, Position) |>
  baretable()
```

## Conference organization

```{r conf-service, results='asis'}
tribble(
  ~Year, ~Details,
  2024, "Program co-chair, Science to Restore and Manage Wildland Aquatic Habitats and Watersheds, AGU Fall Meeting",
  2023, "Workshop leader, Programming with R, Wisconsin Chapter American Fisheries Society Annual Meeting",
  2022, "Workshop leader, Programming with R, Wisconsin Chapter American Fisheries Society Annual Meeting",
  2016, "Program co-chair, Western Division American Fisheries Society Student Colloquium",
  ) |>
  baretable()
```

# University involvement and service

<!-- ## Adjunct and affiliate appointments -->

<!-- ## Graduate student committees -->

<!-- ## Postdoctoral researchers -->

## Invited talks and presentations

* *Freshwater science for tomorrow: advancing fish ecology, conservation, and management*. Wildlife Biology Program, University of Montana, Missoula, Feb 2025 
* *Freshwater science in a changing world: fisheries research for conservation & management*. School of Freshwater Science, UW-Milwaukee, Milwaukee, Dec 2023 
* *Fish and Freshwater in a Changing World: Science for Management and Conservation*. Rocky Mountain Research Station, US Forest Service, Boise, Apr 2023
* *Coldwater Fish in a Changing World: Actionable Ecology for Management and Policy*. Department of Ecology, Montana State University, Bozeman, Mar 2022
* *Exploring isotopic niches in Rocky Mountain Great Plains stream fish*. Dept. of Zoology & Physiology, University of Wyoming, Laramie, May 2017

## Committees, Groups, and Councils

```{r uni-serivce, results='asis'}
tribble(
  ~Start, ~End, ~Position,
  2022, 2024, "Field Operation Facilitation and Boat Committee, Center for Limnology, UW-Madison",
  2019, 2020, "Faculty Search Committee, Department of Zoology and Physiology, University of Wyoming",
  2016, 2020, "Social Media Coordinator, Program in Ecology, University of Wyoming",
  2016, 2017, "Graduate Council, student representative, University of Wyoming",
  2016, 2017, "President, Program in Ecology, University of Wyoming",
  ) |>
  # arrange(-Start) |>
  mutate(
    Start = as.character(Start),
    End = as.character(End),
    End = tidyr::replace_na(End, ""),
    Years = if_else(End == Start,
      as.character(Start),
      paste(Start, "--", End, sep = ""),
      paste(Start, "--", End, sep = "")
    )
  ) |>
  select(Years, Position) |>
  baretable()
```

## Teaching

```{r teaching, results='asis'}
tribble(
  ~ Start, ~ End, ~ Name, ~ Institution,
  2020, 2020, "Instructor, Ichthyology", "University of Wyoming",
  2016, 2019, "Lab Instructor, Ichthyology", "University of Wyoming",
  2015, 2015, "Lab Instructor, Fisheries Management", "University of Wyoming",
  2013, 2014, "Lab Instructor, Fisheries Management", "University of Alberta",
  ) |>
  mutate(
    Start = as.character(Start),
    End = as.character(End),
    End = tidyr::replace_na(End, ""),
    Years = if_else(End == Start,
      as.character(Start),
      paste(Start, "--", End, sep = ""),
      paste(Start, "--", End, sep = "")
    )
  ) |>
  select(Years, Name, Institution) |>
  baretable()
```

## Mentoring

```{r mentoring, results='asis'}
undergrads <- tribble(
  ~ Start, ~ End, ~ Name, ~ Institution,
  2023, 2023, "Jonathan Martin", "University of Wisconsin-Madison",
  2018, 2020, "Nathan Barrus", "University of Wyoming",
  2019, 2020, "Russell Brubaker", "University of Wyoming",
  2018, 2019, "Reilly Davis", "University of Wyoming",
  2017, 2017, "Brynn Hirschman", "University of Wyoming",
  2017, 2017, "Austin Jacobson", "University of Wyoming",
  2016, 2016, "Shaun Helsper", "University of Wyoming",
  2016, 2017, "Travis Allison", "University of Wyoming",
  2016, 2016, "Andrew Annear", "University of Wyoming",
  2015, 2015, "Jake Werner", "University of Wyoming",
  ) |>
  mutate(
    Start = as.character(Start),
    End = as.character(End),
    End = tidyr::replace_na(End, ""),
    Years = if_else(End == Start,
      as.character(Start),
      paste(Start, "--", End, sep = ""),
      paste(Start, "--", End, sep = "")
    )
  ) |>
  select(Years, Name, Institution)

# undergrads |>
#   baretable()
```

* I have mentored `r nrow(undergrads)` undergraduate students as research assistants in lab and field settings at the Univerisity of Wyoming and University of Wisconsin-Madison. 

# Public Service and Outreach

## Public Lectures

* *Climate change, extreme weather, and stream trout in Wisconsin*. Trout Unlimited, Madison, WY, Mar 2022
* *Stream culverts and fish: the good, the bad, and the ugly*. Pinedale Science Café, Pinedale, WY, Aug 2017
* *Culverts as culprits: how road crossings affect fish*. Laramie Science Café, Laramie, WY, Jun 2016 

## Selected Media

* Milwaukee Journal Sentinel, Jan 2023. ["Climate change is making Wisconsin trout more vulnerable to floods and heat"](https://www.jsonline.com/story/news/2023/01/25/climate-change-is-making-wisconsin-trout-more-vulnerable-study-shows/69800081007/). (Also covered in separate stories on [WPR Public Radio](https://www.wpr.org/shows/morning-show/miss-wisconsin-crowned-miss-america-trout-feeling-affects-climate-change))
* Water Blogged, Dec 2022. ["A Tale of Two Fishes: Study Maps the Future of Trout Populations in Wisconsin"](https://blog.limnology.wisc.edu/2022/12/19/a-tale-of-two-fishes-study-maps-the-future-of-trout-populations-in-wisconsin/)
* UW-Madison News, Dec 2022. ["Study maps uneasy future of Wisconsin trout populations"](https://news.wisc.edu/study-maps-uneasy-future-of-wisconsin-trout-populations/)
* The Tributary, Nov 2017. [“Stuck in a rut in Anaconda? Crash a cutthroat spawning party!”](https://wdafs.org/download/Vol-41-No-02.pdf/)

## Service

```{r outreach, results='asis'}
tribble(
  ~ Start, ~ End, ~ Role, ~ Institution,
  2022, 2023, "Food service volunteer", "Ronald McDonald House, Madison, WI",
  2016, 2020, "Science outreach and stream ecology field trips", "Laramie High School, Laramie, WY",
  2016, 2019, "Group leader, Wyoming Bio-Blitz", "Biodiversity Institute, University of Wyoming",
  2016, 2018, "Big Brother", "Big Brothers Big Sisters of Wyoming",
  2017, 2018, "Habitat restoration projects", "Girl Scouts of America",
  2016, 2017, "Science judge", "Wyoming State Science Fair"
  ) |>
  # arrange(-Start) |>
  mutate(
    Start = as.character(Start),
    End = as.character(End),
    End = tidyr::replace_na(End, ""),
    Years = if_else(End == Start,
      as.character(Start),
      paste(Start, "--", End, sep = ""),
      paste(Start, "--", End, sep = "")
    )
  ) |>
  select(Years, Role, Institution) |>
  baretable()
```

# Presentations (chronological)

## Invited

\refstepcounter{bib}\label{presstart}

```{r pres-invite}
bibliography_entries("refs_presentations.bib") |> 
  filter(note == "Invited") |> 
  arrange(issued, desc(author$family))
```

## Offered

```{r pres-offered}
bibliography_entries("refs_presentations.bib") |> 
  filter(note == "Offered") |> 
  arrange(issued, desc(author$family))
```

\refstepcounter{bib}\label{presend}