---
name: Bryan M. Maitland
position: "Research Fish Biologist"
address: "USDA Forest Service,\\tabularnewline Rocky Mountain Research Station, Boise, ID, USA"
email: "bmaitland101@gmail.com"
github: bmait101
date: "`r format(Sys.time(), '%B %Y')`"
output: vitae::hyndman
header-includes:
  \definecolor{headcolor}{HTML}{000088}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  warning = FALSE, 
  message = FALSE
  )
library(vitae)
library(dplyr)
library(RefManageR)
source(here::here("R/pretty.R"))
source(here::here("R/rpackages.R"))
source(here::here("R/gcite.R"))
# source(here::here("R/add_bib_section.R"))
```

# Employment history

```{r employment, results='asis'}
tribble(
  ~Start, ~End, ~Position,
  2023, 2024, "\\textbf{Research Associate},  Department of Civil and Environmental Engineering, University of Wisconsin-Madison",
  2024, NA, "\\textbf{Research Fish Biologist}, Rocky Mountain Research Station, USDA Forest Service",
) |>
  arrange(-Start) |>
  mutate(
    Start = as.character(Start),
    End = as.character(End),
    End = tidyr::replace_na(End, ""),
    Years = if_else(End == Start,
      as.character(Start),
      paste(Start, "--", End, sep = ""),
      paste(Start, "--", End, sep = "")
    )
  ) |>
  select(Years, Position) |>
  baretable()
```

# Education and preparation

```{r degrees, results='asis'}
tribble(
  ~ Degree, ~ Year, ~ Institution,
  "EPA Postdoctoral Fellow", "2022-2023", "University of Wisconsin-Madison",
  "Water Resources Fellow", "2020-2022", "University of Wisconsin-Madison",
  "Ph.D. Ecology", "2020", "University of Wyoming",
  "M.Sc. Conservation Biology", "2015", "University of Alberta",
  "B.S. Conservation Biology", "2011", "SUNY: ESF"
  ) |> 
  baretable()

# tribble(
  # ~ Degree, ~ Year, ~ Institution, ~ Where,
  # "EPA Postdoctoral Fellow", "2022-2023", "University of Wisconsin-Madison", "Madison, WI, USA",
  # "Water Resources Postdoctoral Fellow", "2020-2022", "University of Wisconsin-Madison", "Madison, WI, USA",
  # "Ph.D. Ecology", "2020", "University of Wyoming", "Laramie, WY, USA",
  # "M.Sc. Conservation Biology", "2015", "University of Alberta", "Edmonton, AB, CAN",
  # "B.S. Conservation Biology", "2011", "SUNY: ESF", "Syracuse, NY, USA"
# ) %>% 
#   detailed_entries(Degree, Year, Institution, Where)
```

# Awards and honours

```{r awards, results='asis'}
tribble(
  ~Year, ~Award,
  2024, "Outstanding Service Award, USDA FS Rocky Mountain Research Station",
  2021, "Professional Presentation Award, Wisconsin Chapter American Fisheries Society",
  2019, "Dean’s Graduate Scholar Award, College of Arts and Sciences University of Wyoming",
  2018, "Eugene Maughan Graduate Award, Western Division American Fisheries Society",
  2018, "Hank Gardner Graduate Award, Department of Zoology and Physiology University of Wyoming",
  2017, "Clemens-Rigler Award, Canadian Aquatic Resources Section American Fisheries Society",
  2017, "Board of Visitors Award, College of Arts and Sciences University of Wyoming",
  2017, "Dennis Anderson Fisheries Award, Department Zoology and Physiology University of Wyoming ",
  2017, "William Trachtenberg Scholarship Award, Sustainable Fisheries Foundation",
  2017, "Vern Bressler Fisheries Award, Department of Zoology and Physiology University of Wyoming",
  2015, "Excellence in Fisheries Award, Canadian Aquatic Resources Section American Fisheries Society",
  2015, "Clemens-Rigler Award, Canadian Aquatic Resources Section American Fisheries Society"
  ) |>
  baretable()
```

# Research

```{r}
pubs <- RefManageR::ReadBib("refs_articles.bib", check = FALSE) |> 
  as_tibble() |> 
  filter(bibtype == "Article")
```

  * Since 2015 I have authored `r NROW(pubs)` papers in peer-reviewed journals on fisheries and conservation topics. 
  * On Google Scholar my h-index is `r bmm_cite$All[2]` with total citations of `r prettyNum(bmm_cite$All[1], big.mark=",")` (as at `r format(Sys.time(), "%e %B %Y")`).
  * I have coauthored `r NROW(bmm_pkgs)` R package as a result of my research. There have been over `r round(sum(bmm_pkgs$downloads, na.rm=TRUE))`  downloads of my packages.

# Grants

## Awarded

```{r grants}
grants <- readr::read_csv("grant_income.csv")

sumgrants <- grants |> 
  filter(Status %in% c("Complete", "In progress")) |> 
  pull(Amount) |> 
  sum(na.rm = TRUE)
```

I have acquired `r scales::dollar_format()(sumgrants)` in research grants since 2015.\vspace*{0.2cm}\enlargethispage*{0.5cm}

```{r grantstable, results='asis'}
grants |>
  # filter(Amount > 100000, Type == "Research") |>
  filter(Status %in% c("Complete", "In progress", "Pending")) |> 
  mutate(
    Time = paste0(Start, ifelse(is.na(End), "", paste0("--", End))),
    Entry = paste(People, ". ``", Title, "''. ",
      "\\emph{Funding from ", Source, "}",
      sep = ""
    ),
    Amount = dollars(Amount)
  ) |>
  arrange(desc(Time), desc(Amount)) |>
  select(Time, Entry, Amount) %>%
  baretable(
    align = "llp{13.3cm}r",
    hline.after = c(0, NROW(.)),
    size = "\\fontsize{10}{12}",
    add.to.row = list(
      pos = as.list(seq(NROW(.))),
      command = rep("\\\\[-0.32cm]", NROW(.))
    )
  )
```

\vspace*{0.2cm}\enlargethispage*{0.5cm} 

## Invited, not awarded

Below are invited full proposals that were not awarded: \vspace*{0.2cm}\enlargethispage*{0.5cm}

```{r grantstable-invite, results='asis'}
grants |>
  filter(!Status %in% c("Complete", "In progress", "Pending")) |> 
  mutate(
    Time = paste0(Start, ifelse(is.na(End), "", paste0("--", End))),
    Entry = paste(People, ". ``", Title, "''. ",
      "\\emph{Funding from ", Source, "}",
      sep = ""
    ),
    Amount = dollars(Amount)
  ) |>
  arrange(desc(Time), desc(Amount)) |>
  select(Time, Entry, Amount) %>%
  baretable(
    align = "llp{13.3cm}r",
    hline.after = c(0, NROW(.)),
    size = "\\fontsize{10}{12}",
    add.to.row = list(
      pos = as.list(seq(NROW(.))),
      command = rep("\\\\[-0.32cm]", NROW(.))
    )
  )
```

# Publications

## Theses

```{r}
bibliography_entries("refs_theses.bib") |> 
  arrange(issued)
```

## Pre-prints and in-review

```{r}
bibliography_entries("refs_articles.bib") |> 
  filter(type == "manuscript") |> 
  arrange(issued, desc(author$family))
```

## Refereed research papers

```{r}
bibliography_entries("refs_articles.bib") |> 
  filter(type == "article-journal") |> 
  arrange(issued, desc(author$family))
```

## Technical Reports

```{r}
bibliography_entries("refs_reports.bib") %>%
  arrange(issued, desc(author$family))
```

## Unpublished working papers

TO COME

## Software

```{r}
bibliography_entries("refs_pkgs.bib")
```

# Professional Service

## Ad-hoc reviewer

## Granting agency external reviewer

## Conference organization and serive

## Profressional society service

## Professional society membership

## University service

# Teching

# Mentoring

# Media

# Special skills and training

# Community service and outreach

# Pre-doctoral research and field experience

