---
name: Bryan M. Maitland
position: "Research Fish Biologist"
address: "USDA Forest Service,\\tabularnewline Rocky Mountain Research Station, Boise, ID, USA"
email: "bmaitland101@gmail.com"
github: bmait101
date: "`r format(Sys.time(), '%B %Y')`"
output: vitae::hyndman
header-includes:
  \definecolor{headcolor}{HTML}{000088}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  warning = FALSE, 
  message = FALSE
  )
library(vitae)
library(dplyr)
library(RefManageR)
source(here::here("R/pretty.R"))
source(here::here("R/rpackages.R"))
source(here::here("R/gcite.R"))
# source(here::here("R/add_bib_section.R"))
```

# Employment history

```{r employment, results='asis'}
tribble(
  ~Start, ~End, ~Position,
  2023, 2024, "\\textbf{Research Associate}, Department of Civil and Environmental Engineering, University of Wisconsin-Madison",
  2024, NA, "\\textbf{Research Fish Biologist}, Rocky Mountain Research Station, USDA Forest Service",
) |>
  arrange(-Start) |>
  mutate(
    Start = as.character(Start),
    End = as.character(End),
    End = tidyr::replace_na(End, ""),
    Years = if_else(End == Start,
      as.character(Start),
      paste(Start, "--", End, sep = ""),
      paste(Start, "--", End, sep = "")
    )
  ) |>
  select(Years, Position) |>
  baretable()
```

# Education and preparation

```{r degrees, results='asis'}
tribble(
  ~ Degree, ~ Year, ~ Institution,
  "EPA Postdoctoral Fellow", "2022-2023", "University of Wisconsin-Madison",
  "Water Resources Fellow", "2020-2022", "University of Wisconsin-Madison",
  "Ph.D. Ecology", "2020", "University of Wyoming",
  "M.Sc. Conservation Biology", "2015", "University of Alberta",
  "B.S. Conservation Biology", "2011", "SUNY: ESF"
  ) |> 
  baretable()

# tribble(
  # ~ Degree, ~ Year, ~ Institution, ~ Where,
  # "EPA Postdoctoral Fellow", "2022-2023", "University of Wisconsin-Madison", "Madison, WI, USA",
  # "Water Resources Postdoctoral Fellow", "2020-2022", "University of Wisconsin-Madison", "Madison, WI, USA",
  # "Ph.D. Ecology", "2020", "University of Wyoming", "Laramie, WY, USA",
  # "M.Sc. Conservation Biology", "2015", "University of Alberta", "Edmonton, AB, CAN",
  # "B.S. Conservation Biology", "2011", "SUNY: ESF", "Syracuse, NY, USA"
# ) %>% 
#   detailed_entries(Degree, Year, Institution, Where)
```

# Awards and honours

```{r awards, results='asis'}
tribble(
  ~Year, ~Award,
  2024, "Outstanding Service Award, USDA FS Rocky Mountain Research Station",
  2021, "Professional Presentation Award, Wisconsin Chapter American Fisheries Society",
  2019, "Dean’s Graduate Scholar Award, College of Arts and Sciences University of Wyoming",
  2018, "Eugene Maughan Graduate Award, Western Division American Fisheries Society",
  2018, "Hank Gardner Graduate Award, Department of Zoology and Physiology University of Wyoming",
  2017, "Clemens-Rigler Award, Canadian Aquatic Resources Section American Fisheries Society",
  2017, "Board of Visitors Award, College of Arts and Sciences University of Wyoming",
  2017, "Dennis Anderson Fisheries Award, Department Zoology and Physiology University of Wyoming ",
  2017, "William Trachtenberg Scholarship Award, Sustainable Fisheries Foundation",
  2017, "Vern Bressler Fisheries Award, Department of Zoology and Physiology University of Wyoming",
  2015, "Excellence in Fisheries Award, Canadian Aquatic Resources Section American Fisheries Society",
  2015, "Clemens-Rigler Award, Canadian Aquatic Resources Section American Fisheries Society"
  ) |>
  baretable()
```

# Research

```{r}
pubs <- RefManageR::ReadBib("refs_articles.bib", check = FALSE) |> 
  as_tibble() |> 
  filter(bibtype == "Article")
```

-   Since 2015 I have authored `r NROW(pubs)` papers in peer-reviewed journals on fisheries and conservation topics.
-   On Google Scholar my h-index is `r bmm_cite$All[2]` with total citations of `r prettyNum(bmm_cite$All[1], big.mark=",")` (as at `r format(Sys.time(), "%e %B %Y")`).
-   I have coauthored `r NROW(bmm_pkgs)` R package as a result of my research. There have been over `r round(sum(bmm_pkgs$downloads, na.rm=TRUE))` downloads of my packages.

# Grants

## Awarded

```{r grants}
grants <- readr::read_csv("grant_income.csv")

sumgrants <- grants |> 
  filter(Status %in% c("Complete", "In progress")) |> 
  pull(Amount) |> 
  sum(na.rm = TRUE)
```

I have acquired `r scales::dollar_format()(sumgrants)` in research grants since 2015.\vspace*{0.2cm}\enlargethispage*{0.5cm}

```{r grantstable, results='asis'}
grants |>
  # filter(Amount > 100000, Type == "Research") |>
  filter(Status %in% c("Complete", "In progress", "Pending")) |> 
  mutate(
    Time = paste0(Start, ifelse(is.na(End), "", paste0("--", End))),
    Entry = paste(People, ". ``", Title, "''. ",
      "\\emph{Funding from ", Source, "}",
      sep = ""
    ),
    Amount = dollars(Amount)
  ) |>
  arrange(desc(Time), desc(Amount)) |>
  select(Time, Entry, Amount) %>%
  baretable(
    align = "llp{13.3cm}r",
    hline.after = c(0, NROW(.)),
    size = "\\fontsize{10}{12}",
    add.to.row = list(
      pos = as.list(seq(NROW(.))),
      command = rep("\\\\[-0.32cm]", NROW(.))
    )
  )
```

\vspace*{0.2cm}\enlargethispage*{0.5cm}

## Invited, not awarded

Below are invited full proposals that were not awarded: \vspace*{0.2cm}\enlargethispage*{0.5cm}

```{r grantstable-invite, results='asis'}
grants |>
  filter(!Status %in% c("Complete", "In progress", "Pending")) |> 
  mutate(
    Time = paste0(Start, ifelse(is.na(End), "", paste0("--", End))),
    Entry = paste(People, ". ``", Title, "''. ",
      "\\emph{Funding from ", Source, "}",
      sep = ""
    ),
    Amount = dollars(Amount)
  ) |>
  arrange(desc(Time), desc(Amount)) |>
  select(Time, Entry, Amount) %>%
  baretable(
    align = "llp{13.3cm}r",
    hline.after = c(0, NROW(.)),
    size = "\\fontsize{10}{12}",
    add.to.row = list(
      pos = as.list(seq(NROW(.))),
      command = rep("\\\\[-0.32cm]", NROW(.))
    )
  )
```

# Publications

## Theses

```{r}
bibliography_entries("refs_theses.bib") |> 
  arrange(issued)
```

## Pre-prints and in-review

```{r}
bibliography_entries("refs_articles.bib") |> 
  filter(type == "manuscript") |> 
  arrange(issued, desc(author$family))
```

## Refereed research papers

```{r}
bibliography_entries("refs_articles.bib") |> 
  filter(type == "article-journal") |> 
  arrange(issued, desc(author$family))
```

## Technical Reports

```{r}
bibliography_entries("refs_reports.bib") %>%
  arrange(issued, desc(author$family))
```

## Unpublished working papers

TO COME

## Software

```{r}
bibliography_entries("refs_pkgs.bib")
```

# Professional Service

## Ad-hoc reviewer

```{r reviews1}
n_reviewed <- 49
n_journals <- 15
n_per_yr <- round(n_reviewed / (2025 - 2016), 1)
```

-   I have served as an ad-hoc reviewer for `r n_reviewed` manuscripts from `r n_journals` journals since 2016 ($\sim$ `r n_per_yr` per year) including: *Oecologia*, *Global Change Biology*, *Ecological Applications*, *Freshwater Biology*, *Freshwater Science*, *Frontiers in Ecology and Evolution*, *Hydrobiologia*, *Journal of Ecology*, *Fisheries Research*, *Environmental Biology of Fishes*, *Restoration Ecology*, *PeerJ*, *River Research and Applications*, *Journal of Great Lakes Research*, and *Diversity and Distributions*.

## Other peer review

```{r reviews2, results='asis'}
tribble(
  ~Year, ~Details,
  2025, "Lake Champlain Fisheries Research Program, full proposal review",
  2024, "EPA, internal science review",
  2023, "NSF, Division of Environmental Biology, CAREER, full proposal review",
  2023, "USGS, internal science review",
  ) |>
  baretable()
```

## Conference organization and serive

```{r conf-service, results='asis'}
tribble(
  ~Year, ~Details,
  2024, "Abstract reviewer, American Geophysical Union, Fall Meeting",
  ) |>
  baretable()
```

## Profressional society service

```{r society-service, results='asis'}
tribble(
  ~Start, ~End, ~Position,
  2015, NA, "\\textbf{Member}, Freshwater Working Group, Society for Conservation Biology",
  2021, 2022, "\\textbf{Member}, Fisheries Working Group, Wisconsin Initiative on Climate Change Impacts",
  2020, 2022, "\\textbf{Member}, Trout Species Team, Wisconsin Department of Natural Resources",
  2016, 2017, "\\textbf{President}, UWyo Student Chapter, American Fisheries Society",
  2016, 2018, "\\textbf{Excom member}, CO-WY State Chapter, American Fisheries Society",
  ) |>
  # arrange(-Start) |>
  mutate(
    Start = as.character(Start),
    End = as.character(End),
    End = tidyr::replace_na(End, ""),
    Years = if_else(End == Start,
      as.character(Start),
      paste(Start, "--", End, sep = ""),
      paste(Start, "--", End, sep = "")
    )
  ) |>
  select(Years, Position) |>
  baretable()
```

## Professional society membership

```{r societies, results='asis'}
tribble(
  ~Start, ~End, ~Position,
  2010, NA, "Society for Conservation Biology",
  2010, NA, "American Fisheries Society",
  2015, NA, "Ecological Society of America",
  2016, NA, "Society for Freshwater Science",
  2019, NA, "Galápagos Conservancy",
  2016, 2019, "Association for the Science of Limnology and Oceanography",
  2010, 2016, "International Association of Great Lakes Research",
  ) |>
  # arrange(-Start) |>
  mutate(
    Start = as.character(Start),
    End = as.character(End),
    End = tidyr::replace_na(End, ""),
    Years = if_else(End == Start,
      as.character(Start),
      paste(Start, "--", End, sep = ""),
      paste(Start, "--", End, sep = "")
    )
  ) |>
  select(Years, Position) |>
  baretable()
```

## University service

```{r uni-serivce, results='asis'}
tribble(
  ~Start, ~End, ~Position,
  2022, 2024, "Field Operation Facilitation and Boat Committee, Center for Limnology, UW-Madison",
  2019, 2020, "Faculty Search Committee, Department of Zoology and Physiology, University of Wyoming",
  2016, 2020, "Social Media Coordinator, Program in Ecology, University of Wyoming",
  2016, 2017, "Graduate Council, student representative, University of Wyoming",
  2016, 2017, "President, Program in Ecology, University of Wyoming",
  ) |>
  # arrange(-Start) |>
  mutate(
    Start = as.character(Start),
    End = as.character(End),
    End = tidyr::replace_na(End, ""),
    Years = if_else(End == Start,
      as.character(Start),
      paste(Start, "--", End, sep = ""),
      paste(Start, "--", End, sep = "")
    )
  ) |>
  select(Years, Position) |>
  baretable()
```

# Teching

```{r teaching, results='asis'}
tribble(
  ~ Year, ~ Course, ~ Role, ~ Institution,
  "2020", "ZOO 4330 - Ichthyology", "Instructor", "University of Wyoming",
  "2019", "ZOO 4330 - Ichthyology", "Lab Instructor", "University of Wyoming",
  "2018", "ZOO 4330 - Ichthyology", "Lab Instructor", "University of Wyoming",
  "2017", "ZOO 4330 - Ichthyology", "Lab Instructor", "University of Wyoming",
  "2016", "ZOO 4330 - Ichthyology", "Lab Instructor", "University of Wyoming",
  "2015", "ZOO 4310 - Fisheries Management", "Lab Instructor", "University of Wyoming",
  "2014", "RENR 460 - Fisheries Management", "Lab Instructor", "University of Alberta",
  "2013", "RENR 460 - Fisheries Management", "Lab Instructor", "University of Alberta",
  ) |> 
  baretable()
```

# Mentoring

## Undergraduate Students (Research Assistants, lab and field)

```{r mentoring, results='asis'}
tribble(
  ~ Start, ~ End, ~ Name, ~ Institution,
  2023, 2023, "Jonathan Martin", "University of Wisconsin-Madison",
  2018, 2020, "Nathan Barrus", "University of Wyoming",
  2019, 2020, "Russell Brubaker", "University of Wyoming",
  2018, 2019, "Reilly Davis", "University of Wyoming",
  2017, 2017, "Brynn Hirschman", "University of Wyoming",
  2017, 2017, "Austin Jacobson", "University of Wyoming",
  2016, 2016, "Shaun Helsper", "University of Wyoming",
  2016, 2017, "Travis Allison", "University of Wyoming",
  2016, 2016, "Andrew Annear", "University of Wyoming",
  2015, 2015, "Jake Werner", "University of Wyoming",
  ) |>
  baretable()
```

# Media


# Special skills and training

# Community service and outreach

# Pre-doctoral research and field experience
