---
name: Bryan M. Maitland, Ph.D.
position: "Research Fish Biologist"
docname: "Curriculum Vitae"
address: "Rocky Mountain Research Station,\\tabularnewline USDA Forest Service, Boise, ID, USA"
email: "bmaitland101@gmail.com"
orcid: 0000-0002-4491-5064
linkedin: bryanmaitland
github: bmait101
date: "`r format(Sys.time(), '%B %Y')`"
date-format: "MMMM YYYY"
csl: "csl/apa_cv.csl"
format: cv-pdf
filters:
  - lua/names.lua
header-includes:
  \definecolor{headcolor}{HTML}{005838}
---

\newcounter{bib}

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
  )

library(vitae)
library(dplyr)
library(RefManageR)
source(here::here("R/pretty.R"))
source(here::here("R/bib_section.R"))

# Make sure RSTUDIO_PANDOC is set before anything below (bibliography_entries()
# included) needs pandoc - see find_pandoc()'s comment in R/bib_section.R.
find_pandoc()

# Remove any temporary bib files from a previous render, then recreate
# temp.bib empty so the cv-pdf template's unconditional
# \addbibresource{temp.bib} doesn't fail to find the file. The selected-pubs
# file is per-document (see selected_bib below) so that this and the 1-pager
# never clobber each other's subset.
if (fs::file_exists(here::here("temp.bib"))) {
  fs::file_delete(here::here("temp.bib"))
}
fs::file_create(here::here("temp.bib"))

selected_bib <- "selected_pubs_2page.bib"
if (fs::file_exists(here::here(selected_bib))) {
  fs::file_delete(here::here(selected_bib))
}

targets::tar_load(c(bmm_cite, bmm_cite_papers, bmm_pkgs))
```

```{r tuning, include=FALSE}
# ---------------------------------------------------------------------------
# TUNE PER PROPOSAL ---------------------------------------------------------
#
# Everything a given grant proposal is likely to need changed lives either
# here or in the "Research summary" prose immediately below. The rest of the
# document builds itself from the same bibs/ and *.csv sources as the full CV,
# so it stays in sync automatically.
#
# 1. Which papers to feature. Keys are from bibs/journal_articles.bib; they
#    print top-to-bottom in the order listed here, so lead with the ones most
#    relevant to the call. At 8 papers and 7 grants the document ends about a
#    quarter of the way down page 2, which is deliberate slack: it absorbs a
#    longer research summary, or two or three more papers, without spilling
#    onto a third page.
selected_pub_keys <- c(
  "Sparks2026a",   # hatchR: developmental phenology toolset
  "Maitland2024",  # Lake Michigan habitat coupling (top-cited, Ecology)
  "Maitland2022",  # climate + long-term monitoring -> stream trout recruitment
  "Kirk2022",      # climatic drivers and impacts of a rapid range expansion
  "Maitland2023",  # food web expansion along the RM-Great Plains ecotone
  "Barrus2024",    # mentee-led; isotope baselines in streams
  "Maitland2016a", # land-use disturbance -> instream habitat + fish community
  "Craig2017"      # interacting threats in freshwater ecosystems
)

# 2. How many funded/pending grants to list, and the floor for inclusion.
n_grants_shown <- 7
grant_floor <- 20000

# 3. Set to TRUE to include the withdrawn/declined invited full proposals
#    (useful where a sponsor wants evidence of large-proposal leadership).
#    This block costs roughly a third of a page, so drop selected_pub_keys to
#    about 5 and n_grants_shown to about 5 to stay within two pages.
show_invited_proposals <- FALSE
# ---------------------------------------------------------------------------

pubs <- RefManageR::ReadBib("bibs/journal_articles.bib", check = FALSE)

missing_keys <- setdiff(selected_pub_keys, names(pubs))
if (length(missing_keys) > 0) {
  stop(
    "selected_pub_keys not found in bibs/journal_articles.bib: ",
    paste(missing_keys, collapse = ", ")
  )
}

n_pubs <- pubs |>
  as_tibble() |>
  filter(bibtype == "Article") |>
  nrow()

n_talks <- RefManageR::ReadBib("bibs/conf_presentations.bib", check = FALSE) |>
  as_tibble() |>
  nrow()

grants <- readr::read_csv("grant_income.csv")

sumgrants <- grants |>
  filter(Status %in% c("Complete", "In progress")) |>
  pull(Amount) |>
  sum(na.rm = TRUE)
```

# Appointments

```{r employment, results='asis'}
tribble(
  ~Start, ~End, ~Position,
  2024, NA, "\\textbf{Research Fish Biologist}, Rocky Mountain Research Station, USDA Forest Service",
  2023, 2024, "\\textbf{Research Associate}, Dept. of Civil and Environmental Engineering, University of Wisconsin-Madison",
  2022, 2023, "\\textbf{Postdoctoral Fellow}, Center for Limnology, University of Wisconsin-Madison",
  2020, 2022, "\\textbf{WI Water Resources Postdoctoral Fellow}, Aquatic Sciences Center (UW-Madison) and WDNR"
  ) |>
  arrange(-Start) |>
  mutate(
    Start = as.character(Start),
    End = as.character(End),
    End = tidyr::replace_na(End, "Present"),
    Years = paste(Start, "--", End, sep = "")
  ) |>
  select(Years, Position) |>
  baretable()
```

Affiliate: Graduate Faculty, Dept. of Biology, Geology, and Environmental Science, University of Tennessee at Chattanooga (2026--present).

# Education

```{r degrees, results='asis'}
tribble(
  ~ Year, ~ Degree, ~ Feild, ~ Institution,
  2020, "Ph.D.", "Ecology", "University of Wyoming",
  2015, "M.Sc.", "Conservation Biology (Fisheries)", "University of Alberta",
  2011, "B.S.", "Conservation Biology \\textit{cum laude}", "SUNY: College of Environmental Science and Forestry",
  ) |>
  baretable()
```

# Research summary

<!-- TUNE PER PROPOSAL: rewrite the paragraph below to foreground the work most
     relevant to the call. The bullets underneath are generated from the same
     sources as the full CV and normally need no editing.

     CURRENTLY TARGETED AT: JFSP FLAMES 2027 (FA-NOFO0027-001) - Ferone (student
     investigator) / Caudill (PI), wildfire exposure metrics and benthic
     macroinvertebrates in Idaho streams. The generic version of this paragraph
     is in git history at commit 01b17ff. -->

I am a conservation ecologist at the Rocky Mountain Research Station studying how
disturbance---wildfire, climate change, and altered flow regimes---propagates
through stream ecosystems and the food webs they support. I combine stable isotope
ecology, otolith and tissue chemistry, long-term agency biomonitoring records,
remote sensing, and spatial and Bayesian modeling; much of my published work
integrates multi-decade state and federal monitoring datasets with spatially
explicit habitat and climate data to separate direct from indirect drivers of
community change. I am a partner on the JFSP-funded project *From flames to fish:
co-management for wildfire and aquatic species in Idaho* (2026--2028), I co-chaired
the AGU session *Science to Restore and Manage Wildland Aquatic Habitats and
Watersheds* (2024), and I co-chair the *Fish and Fire* symposium at the 2026 AFS
Annual Meeting. A through-line in my work is building open, reusable tools and
datasets---R packages, databases, and reproducible workflows---and teaching R
workshops so that the methods transfer to the practitioners who must act on the
results. I mentor graduate and undergraduate researchers, including Garrison
Ferone at the University of Idaho.

<!-- END TUNE PER PROPOSAL -->

- Since 2014 I have authored `r n_pubs` peer-reviewed journal articles and `r NROW(bmm_pkgs)` R package (`r round(sum(bmm_pkgs$downloads, na.rm=TRUE))` downloads), and presented `r n_talks` papers and posters at AFS, SFS, ASLO, AGU, IAGLR, and regional meetings.
- Google Scholar h-index `r bmm_cite$All[2]`, `r prettyNum(bmm_cite$All[1], big.mark=",")` total citations (as of `r format(Sys.time(), "%e %B %Y")`).
- Since 2015 I have led, co-led, or partnered on \$`r sprintf("%2.1f", sumgrants/1e6)` million in funded projects, and contributed to invited proposals exceeding \$2 million in scope.
- I have reviewed 60 manuscripts for 19 journals since 2016, and served as a grant reviewer for NSF (Division of Environmental Biology) and the Lake Champlain Fisheries Research Program.

# Selected publications

\textbf{Bold} indicates my name; \underline{underlined names} indicate undergraduate research mentees. Citation counts are from Google Scholar on `r format(Sys.time(), "%e %B %Y")`. A full list of `r n_pubs` articles is available in my complete CV.

```{r selected-pubs, results='asis'}
# The CSL sorts the bibliography by citation-number descending, i.e. entries
# print in the reverse of the order they are cited in. Reverse the keys here
# so that selected_pub_keys reads in the same order it renders on the page.
selected <- pubs[rev(selected_pub_keys)]
selected_cites <- get_citation_counts(selected, bmm_cite_papers)

# bibliography_entries() only accepts a bib *file path*, not a BibEntry
# object, so write the selected subset out to its own small bib file first.
RefManageR::WriteBib(selected, file = here::here(selected_bib))

bibliography_entries(selected_bib) |>
  render_bib_section(selected_bib, cites = selected_cites)
```

# Selected funding

```{r grants-table, results='asis'}
grants |>
  filter(Status %in% c("Complete", "In progress", "Pending")) |>
  filter(Amount >= grant_floor) |>
  mutate(
    Time = paste0(Start, ifelse(is.na(End), "", paste0("--", End))),
    Entry = paste(
      Role, ". ``", Title, "''. ",
      "\\emph{", Source, "}.",
      sep = ""
    ),
    # Lead with active funded work, then the pending pipeline, then closed
    # awards; within each group, most recent and largest first. Without this
    # a short list fills up with nothing but "Pending" rows.
    status_rank = match(Status, c("In progress", "Pending", "Complete")),
    sort_year = suppressWarnings(as.numeric(Start)),
    sort_year = tidyr::replace_na(sort_year, Inf)
  ) |>
  arrange(status_rank, desc(sort_year), desc(Amount)) |>
  head(n_grants_shown) |>
  mutate(Amount = dollars(Amount)) |>
  select(Time, Entry, Amount) %>%
  baretable(
    align = "llp{12.8cm}r",
    size = "\\fontsize{9}{11}",
    add.to.row = list(
      pos = as.list(seq(NROW(.))),
      command = rep("\\\\[-0.30cm]", NROW(.))
    )
  )
```

```{r grants-invited, results='asis'}
# The toggle is applied inside the chunk rather than as `eval=` in the chunk
# header. knitr evaluates chunk options in an empty environment when purling,
# and tarchetypes::tar_quarto() purls the .qmd to find tar_load() calls, so a
# variable in a chunk option raises "object not found" on every tar_make().
if (show_invited_proposals) {
  cat("\n\\vspace*{0.15cm}\n\n**Invited full proposals (withdrawn due to U.S. federal funding disruptions, or not awarded):**\n\n\\vspace*{0.15cm}\n\n")

  grants |>
    filter(Status %in% c("Invited, withdrawn", "Invited, not awarded")) |>
    mutate(
      Time = paste0(Start, ifelse(is.na(End), "", paste0("--", End))),
      Entry = paste(
        Role, ". ``", Title, "''. ",
        "\\emph{", Source, "}.",
        sep = ""
      ),
      sort_year = suppressWarnings(as.numeric(Start)),
      sort_year = tidyr::replace_na(sort_year, Inf)
    ) |>
    arrange(desc(sort_year), desc(Amount)) |>
    mutate(Amount = dollars(Amount)) |>
    select(Time, Entry, Amount) %>%
    baretable(
      align = "llp{12.8cm}r",
      size = "\\fontsize{9}{11}",
      add.to.row = list(
        pos = as.list(seq(NROW(.))),
        command = rep("\\\\[-0.30cm]", NROW(.))
      )
    )
}
```

# Selected awards and honors

```{r awards, results='asis'}
tribble(
  ~Year, ~Award,
  2026, "Outstanding Service Award, USDA Forest Service, Rocky Mountain Research Station",
  2024, "Outstanding Service Award, USDA Forest Service, Rocky Mountain Research Station",
  2021, "Best Professional Presentation Award, Wisconsin Chapter American Fisheries Society",
  2020, "Wisconsin Water Resources Postdoctoral Fellow, University of Wisconsin-Madison",
  2019, "Dean's Graduate Fellowship, College of Arts and Sciences, University of Wyoming",
  2018, "Eugene Maughan Graduate Award, Western Division American Fisheries Society",
  ) |>
  baretable()
```

# Mentoring and teaching

```{r mentoring-counts, results='asis'}
tribble(
  ~Years, ~Detail,
  "2023--present", "Supervised 3 graduate/post-baccalaureate researchers (U. Idaho, UW-Madison, UT-Chattanooga); MS committee member, UT-Chattanooga",
  "2015--2025", "Mentored 11 undergraduate researchers, two of whom have led or co-authored peer-reviewed publications",
  "2014--2020", "Instructor of record for Ichthyology (UWyo); teaching assistant for 12 sections of Ichthyology and Fisheries Management (UWyo, U. Alberta)",
  ) |>
  baretable(align = "lp{3.2cm}p{13.1cm}")
```

# Selected service and synergistic activities

```{r service, results='asis'}
tribble(
  ~Years, ~Detail,
  "2025--2029", "Mentor of Practice, NSF IRES: Climate Change Resilience in an Aquatic Bright Spot (Jensen, PI, UW-Madison)",
  "2026", "Program co-chair, \\emph{Fish and Fire} symposium, American Fisheries Society Annual Meeting",
  "2024", "Program co-chair, \\emph{Science to Restore and Manage Wildland Aquatic Habitats and Watersheds}, AGU Fall Meeting",
  "2020--2022", "Member, Trout Species Team, Wisconsin Dept. of Natural Resources; Fisheries Working Group, Wisconsin Initiative on Climate Change Impacts",
  "2015--present", "Member, Freshwater Working Group, Society for Conservation Biology; AFS, SCB, ESA, SFS",
  ) |>
  baretable(align = "lp{3.2cm}p{13.1cm}")
```

\vspace*{0.2cm}

Developer and maintainer of `hatchR`, an open-source R package for predicting fish hatch and emergence timing from thermal regimes. Instructor for R programming and data visualization workshops at American Fisheries Society chapter meetings. Research covered by regional press and public radio.
